Topological Equivariant Artist Model for Visualization Library Architecture

### 2022-04-08 paper/main.tex
flipping arrows

\subsection{Data: Base $\rightarrow$ Fiber}
As described in \autoref{sec:atct:fiber-bundles}, we model data as a map from base space to fiber space.To encapsulate the structure carried in these maps, and provide a way to define these maps in reusable and composable manner, we define categories of base and fiber spaces.

\begin{definition}
  The base space category $\mathcal{\opensetc}$ has open set objects $\opensetc$ and inclusion morphisms $\iota: \opensetc_1 \rightarrow \opensetc_2$ such that $\opensetc_1 \subseteq \opensetc_2$. 
\end{definition}

@@

\begin{definition}
  The fiber category $\mathcal{\dfiberc}$ is a monoidal category, meaning the category has a single object $\dfiberc$ of an arbitrary type. The morphisms on the fiber object are $\dfuncc \in Hom(\dfiberc, \dfiberc)$. 
\end{definition}

@@

\subsection{Sheaves}
Sheaves on bundles are an "algebraic data structure", as described by Ghrist \cite{ghristElementaryAppliedTopology2014}, for expressing the bookkeeping involved in subsetting data and composing data that is distributed or streaming. Formally, pre-sheaves are functors, meaning they are functions from objects of one category to objects of another category \cite{WhatFunctorDefinitions}. In this work we consider the presheaf
\begin{equation}
  \label{eq:atct:presheaf}
  \begin{tikzcd}
    \cgamma{\opensetc_1}{\dtotalc\restriction_{\opensetc_1}}  &  & \cgamma{\opensetc_2}{\dtotalc\restriction_{\opensetc_2}} 
    \arrow[ll, "\iota^*"', color=set] \\
    & & \\
    \opensetc_1 
    \arrow[rr, "\iota", hook, color=base] 
    \arrow[uu, "{\sheafc_{\dbasec, \dtotalc}}", maps to, color=sheaf] &  & \opensetc_2 
    \arrow[uu, "{\sheafc_{\dbasec, \dtotalc}}"', maps to, color=sheaf]              
    \end{tikzcd}
\end{equation}

@@

\subsubsection{Data Transformations $\dfuncc=(\dfunchc, \dfunctc)$}
We express transformations of data, such as scaling and rotation, as function on sheaves: 
\note{may not need the diagram here}
\begin{equation}
  \label{eq:atct:sheaves:morphism}
  \begin{tikzcd}
    \cgamma{\opensetc}{\dtotalc\restriction_{\opensetc}} 
    \arrow[rr, "\dfuncpullc", color=action] &  & \cgamma{\opensetc^{\prime}}{\dfuncpullc\dtotalc\restriction_{\opensetc^{\prime}}} & \cgamma{\opensetc^{\prime}}{\dfuncpullc\dtotalc\restriction_{\opensetc^{\prime}}} 
    \arrow[dd, "\dfunctc", color=action] \\
     &  & &       \\
    \opensetc 
    \arrow[uu, maps to,color=sheaf]  &  & \opensetc^{\prime} 
    \arrow[ll, "\dfunchc"', color=action] 
    \arrow[uu, maps to, color=sheaf]                    & \cgamma{\opensetc^{\prime}}{\dfuncpullc\dtotalc\restriction_{\opensetc^{\prime}}}                       
    \end{tikzcd}
\end{equation}

@@

The fiber transformation $\dfunct: \dfuncpull \dtotal_{\dbasepoint^{\prime}} \rightarrow \dfuncpull \dtotal_{\dbasepoint^{\prime}}$ is a morphism on the fiber $\dfunct \in Hom(\dfuncpull\dfiber\restriction_{\dbasepoint^{\prime}},\dfuncpull\dfiber\restriction_{\dbasepoint^{\prime}})$ restricted to a point $\dbasepoint^{\prime} \in \openset^{\prime}$. The fiber transformation is a change in section 
\begin{equation}
  \dfunct: \dfuncpull \dsection \restriction_{\openset} \mapsto \dfuncpull \dsection^{\prime} \restriction_{\openset}
\end{equation}
where $\dsection, \dsection^{\prime} \in \Gamma(\openset^{\prime}, \dfuncpull\dtotal\restriction_{\openset^{\prime}})$. Examples of a fiber transformation are actions on fibers, such as the permutation, ordering, translation, and scaling codified as Steven's measurement scales \cite{stevensTheoryScalesMeasurement1946}.

@@

\subsection{ Data Index \dbasec\ \& Visualization Region \gbasec}

\begin{equation}
  \begin{tikzcd}[row sep=huge]
    \cgamma{\opensetgc}{\gtotalc\restriction_{\opensetgc}} 
    \arrow[rr, "\vindexpushc", color=functor]   &  & \cgamma{\opensetc}{\vindexpushc\gtotalc\restriction_{\opensetc}}\\
    \opensetgc \arrow[rr, "\vindexc", color=functor] 
    \arrow[u, "{\sheafc_{\gbasec, \gtotalc}}", maps to, color=sheaf] 
    \arrow[d, "{\sheafc_{\gbasec, \vindexpullc\dtotalc}}"', dashed, maps to, color=sheaf] &  & \opensetc 
    \arrow[d, "{\sheafc_{\dbasec,\dtotalc}}", maps to, color=sheaf] 
    \arrow[u, "{\sheafc_{\dbasec, \vindexpushc \gtotalc}}"', dashed, maps to, color=sheaf] \\
    \cgamma{\opensetgc}{\vindexpullc\dtotalc\restriction_{\opensetgc}}  &  & \cgamma{\opensetc}{\dtotalc\restriction_{\opensetc}}
    \arrow[ll, "\vindexpullc"', color=functor]      
    \end{tikzcd}
\end{equation}

\section{Artist: Data $\rightarrow$ Graphic}
Because $Hom_{\gbase}(\sheaf_{\gbase, \vindexpull\dtotal}, \sheaf_{\gbase,\gtotal}) = Hom_{\dbase}(\sheaf_{\dbase, \dtotal}, \sheaf_{\dbase, \vindexpush\gtotal})$

\begin{equation}
\begin{tikzcd}
  {\sheafc_{\gbasec,\gtotalc}} 
  \arrow[rr, "\vindexpushc", color=functor]  
  &  & {\sheafc_{\dbasec,\vindexpushc\gtotalc}}   \\
   &  &             \\
  {\sheafc_{\gbasec,\vindexpullc\dtotalc}} 
  \arrow[uu, "\vartistc^{\gbasec}", Rightarrow, color=artist] &  & {\sheafc_{\dbasec,\dtotalc}} 
  \arrow[ll, "\vindexpullc"', color=functor]
   \arrow[uu, "\vartistc^{\dbasec}"', Rightarrow, color=artist] 
   \arrow[lluu, "\vartistc" description, color=artist]
  \end{tikzcd}
\end{equation}

@@

  \note{introduce \phi_M as basis of equivariance}
\section{Discussion}
\subsection{Limitations}
\subsection{future work}

>>>>

\subsection{Input and Output Types}
\label{sec:atct:io}
As described in \autoref{sec:atct:fiber-bundles}, we model data as a map from base space to fiber space.To encapsulate the structure carried in these maps, and provide a way to define these maps in reusable and composable manner, we define categories of base and fiber spaces.

\begin{definition}
  \label{sec:atct:io:base}
  The base space category $\mathcal{\opensetc}$ has open set objects $\opensetc$ and inclusion morphisms $\iota: \opensetc_1 \rightarrow \opensetc_2$ such that $\opensetc_1 \subseteq \opensetc_2$. 
\end{definition}

@@

\begin{definition}
  \label{def:atct:io:fiber}
  The fiber category $\mathcal{\dfiberc}$ is a monoidal category, meaning the category has a single object $\dfiberc$ of an arbitrary type. The morphisms on the fiber object are $\dfunctc \in Hom(\dfiberc, \dfiberc)$. 
\end{definition}

@@

\subsection{Sheaves}
Sheaves on bundles are an "algebraic data structure", as described by Ghrist \cite{ghristElementaryAppliedTopology2014}, for expressing the bookkeeping involved in subsetting data and composing data that is distributed or streaming. Formally, pre-sheaves are functors, meaning they are functions from objects of one category to objects of another category \cite{WhatFunctorDefinitions}. Here the sheaf 
\begin{equation}
  \sheafc_{\dbasec, \dtotalc}: \opensetc \rightarrow \cgamma{\opensetc}{\dtotalc\restriction_{\opensetc}}
\end{equation}
denotes a sheaf from opensets on the base space $\dbase$ to sets of sections on the bundle $\dtotal$. As a contravariant functor it preserves morphism s

\begin{equation}
  \label{eq:atct:presheaf}
  \begin{tikzcd}
    \cgamma{\opensetc_1}{\dtotalc\restriction_{\opensetc_1}}  &  & \cgamma{\opensetc_2}{\dtotalc\restriction_{\opensetc_2}} 
    \arrow[ll, "\iota^*"', color=set] \\
    & & \\
    \opensetc_1 
    \arrow[rr, "\iota", hook, color=base] 
    \arrow[uu, "{\sheafc_{\dbasec, \dtotalc}}", maps to, color=sheaf] &  & \opensetc_2 
    \arrow[uu, "{\sheafc_{\dbasec, \dtotalc}}"', maps to, color=sheaf]              
    \end{tikzcd}
\end{equation}

@@

\subsubsection{Measurement Types}

We specify that structure on a sheaf includes the actions $\dfuncc$ on sections of that sheaf, which is a generalization of Steven's \cite{stevensTheoryScalesMeasurement1946} classification of measurement scales by their mathematical group structure. These sheaf actions describe which changes in the data we expect to be equivariant to changes in the visualization.

\begin{equation}
  \label{eq:atct:sheaves:morphism}
  \begin{tikzcd}
    \cgamma{\opensetc}{\dtotalc\restriction_{\opensetc}} 
    \arrow[rr, "\dfuncpullc", color=action, maps to] &  & 
    \cgamma{\opensetc^{\prime}}{\dfuncpullc\dtotalc\restriction_{\opensetc^{\prime}}} & 
    \cgamma{\opensetc^{\prime}}{\dfuncpullc\dtotalc\restriction_{\opensetc^{\prime}}} 
    \arrow[dd, "\dfunctc", color=action] \\
     &  & &       \\
    \opensetc 
    \arrow[uu, maps to,color=sheaf]  &  & \opensetc^{\prime} 
    \arrow[ll, "\dfunchc"', color=action, maps to] 
    \arrow[uu, maps to, color=sheaf]                    & \cgamma{\opensetc^{\prime}}{\dfuncpullc\dtotalc\restriction_{\opensetc^{\prime}}}                       
    \end{tikzcd}
\end{equation}

@@

As introduced in \autoref{def:atct:io:fiber}, the fiber transformation $\dfunct: \dfuncpull \dtotal_{\dbasepoint^{\prime}} \rightarrow \dfuncpull \dtotal_{\dbasepoint^{\prime}}$ is a morphism on the fiber $\dfunct \in Hom(\dfuncpull\dfiber\restriction_{\dbasepoint^{\prime}},\dfuncpull\dfiber\restriction_{\dbasepoint^{\prime}})$ restricted to a point $\dbasepoint^{\prime} \in \openset^{\prime}$. The fiber transformation is a change in section 
\begin{equation}
  \dfunct: \dfuncpull \dsection \restriction_{\openset} \mapsto \dfuncpull \dsection^{\prime} \restriction_{\openset}
\end{equation}
where $\dsection, \dsection^{\prime} \in \Gamma(\openset^{\prime}, \dfuncpull\dtotal\restriction_{\openset^{\prime}})$. Examples of a fiber transformation are actions on fibers, such as the permutation, ordering, translation, and scaling codified as Steven's measurement scales \cite{stevensTheoryScalesMeasurement1946}.

@@

\subsection{Map Between Base Spaces}
\begin{figure}[ht]
  \includegraphics*[width=1\columnwidth]{../../slides/figures/math/push_pull_scatter.png}
\end{figure}
We define a functor
\begin{equation}
  \vindexc: \opensetgc \rightarrow \openset 
\end{equation}
that expresses the constraint that every region in a visualization $\opensetg \subseteq \gbase$ must map to a corresponding set $\openset \subset \dbase$. We define the functor $\vindexc$ such that the data base space \dbasec\ is a deformation retraction\cite{nlab:deformation_retract} of the graphic space $\gbasec$. To simplify matters, in this paper, we construct the graphic space as a constant multiple of the base space such that 
\begin{equation}
  \vindexc: \underbrace{\opensetc\times[0,1]^{n}}_{\opensetgc} \rightarrow \opensetc 
\end{equation}
where $ n = dim(\dbase) - dim(\gbase)$ if $dim(\dbase)<dim(\gbase)$ and otherwise $ n = dim(\dbase)$. 
\begin{equation}
  \label{eq:atct:morphisms:xi}
\begin{tikzcd}[row sep=huge]
  \cgamma{\opensetc}{\dtotalc\restriction_{\opensetc}} 
  \arrow[rr, "\vindexpullc", maps to, color=functor] &  & 
  \cgamma{\opensetgc}{\vindexpullc\dtotalc\restriction_{\opensetgc}}                                                                                                       \\
  \opensetc 
  \arrow[d, "{\sheafc_{\dbasec, \vindexpushc \gtotalc}}", dashed, maps to, color=sheaf] 
  \arrow[u, "{\sheafc_{\dbasec, \dtotalc}}"', maps to, color=sheaf] &  & 
  \opensetgc 
  \arrow[d, "{\sheafc_{\gbasec, \gtotalc}}", maps to, color=sheaf] 
  \arrow[u, "{\sheafc_{\gbasec, \vindexpullc\dtotalc}}"', dashed, maps to, color=sheaf] 
  \arrow[ll, "\vindexc"', maps to, color=functor] \\
  \cgamma{\opensetc}{\vindexpushc\gtotalc\restriction_{\opensetc}} &  & 
  \cgamma{\opensetgc}{\gtotalc\restriction_{\opensetgc}} 
  \arrow[ll, "\vindexpushc"', maps to, color=functor]                                                         
  \end{tikzcd}
\end{equation}

\section{Artist: Data to Graphic}
Because $Hom_{\gbase}(\sheaf_{\gbase, \vindexpull\dtotal}, \sheaf_{\gbase,\gtotal}) = Hom_{\dbase}(\sheaf_{\dbase, \dtotal}, \sheaf_{\dbase, \vindexpush\gtotal})$

\begin{equation}
  \begin{tikzcd}
    \cgamma{\opensetc}{\dtotalc\restriction_{\opensetc}} 
    \arrow[dd, "\vartistc^{\dbasec}"', color=artist] 
    \arrow[rrdd, "\vartistc", color=artist] 
    \arrow[rr, "\vindexpullc", color=functor] &  &
    \cgamma{\opensetgc}{\vindexpullc\dtotalc\restriction_{\opensetgc}} 
    \arrow[dd, "\vartistc^{\gbasec}", color=artist] \\
     & & \\
    \cgamma{\opensetgc}{\gtotalc\restriction_{\opensetgc}} &  & 
    \cgamma{\opensetgc}{\gtotalc\restriction_{\opensetgc}} 
    \arrow[ll, "\vindexpushc"', color=functor]                  
    \end{tikzcd}
\end{equation}

@@

  \note{introduce $\phi_M$ as basis of equivariance}
\section{Discussion}
\subsection{Limitations}
\subsection{future work}
